EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-20 03_edipack2.tex
Co-authored-by: Samuele Giuli <samuele.giuli@gmail.com>

\subsubsection{Electron-Phonon coupling}
Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_q \omega_{0q} b_q^+b_q\\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, and $b_q$
($b_q^+$) are the destruction (creation) operators for a phonon with
frequency $\omega_{0q}$.
The matrix  $g_{\a\b}$ describe the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$.

>>>>

\subsubsection{Electron-Phonon coupling}
Finally, the electron-phonon part of the quantum impurity problems is
described by the Hamiltonian terms: 
\begin{equation}\label{Hph}
  \begin{split}
    \hat{H}_\mathrm{ph}&=\sum_q \omega_{0q} b_q^+b_q + \sum_q \sqrt{2}A_{q}(b^\dagger_q + b_q)  \\
    % 
    \hat{H}_\mathrm{e-ph} &= \sum_q\sum_{\a\b\sigma} g_{\a\b} d^+_{\a\sigma}d_{\b\sigma}(b_q+b_q^+)
\end{split}
\end{equation}
where $q=1,\dots,N_q$ indexes the number of local phonons, and $b_q$
($b_q^+$) are the destruction (creation) operators for a phonon with
frequency $\omega_{0q}$.
The matrix  $g_{\a\b}$ describe the electron-phonon coupling. 
Although feasible, dealing with more than one phonon mode becomes
quickly computationally very demanding, thus in the rest of the this
work we shall consider $N_q=1$.
The phonon's frequency is passed to the \NAME program via the variable { \tt W0\_PH } to be read from the input file. The displacement field acting on the phonon is passed through the variable {\tt A\_PH} from the input file. The electron-phonon coupling matrix $g_{\alpha \beta}$ instead can be passed in two different ways. If the matrix is diagonal ($g_{\alpha \beta} = g_\alpha \delta_{\alpha \beta}$), from the input file one can simply write in order the values $g_\alpha$ in the variable {\tt G\_PH }. For any general shape one can pass to the input file a {\tt GPHfile } variable containing the name of the file storing the $g_{\alpha \beta}$ array in Fortran style.
If {\tt GPHfile=NONE} the matrix $g_{\alpha \beta}$ is read from the variable {\tt G\_PH}, else it is read from the {\tt GPHfile}.

### 2025-05-20 05_examples.tex
Co-authored-by: Samuele Giuli <samuele.giuli@gmail.com>

This interaction in two limits takes two simpler forms, for $\omega_0 \rightarrow \infty$ with $g^2/\omega_0=\lambda$ constant, also known as anti-adiabatic limit, the interaction becomes a Hubbard attraction:
\begin{equation}
    H_{int} \overset{ \omega_0 \rightarrow \infty}{ \longrightarrow } -\frac{\lambda}{2} \sum_i \Big[\sum_\sigma(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2}) \Big]^2
\end{equation}

Meanwhile in the opposite limit, $\omega_0 \rightarrow 0$ with $g^2/\omega_0=\lambda$ constant, also known as adiabatic limit, the phonons are condensed and the creation/annihilation operators can be considered as a scalar number to be determined self-consistently. In that case the interaction at equation \eqref{eqex:H_Holstein} for uniform solution vanishes and one recovers the non-interacting limit.

THE FOLLOWING MAYBE IN THE E-PH SECTION

The phonon's frequency is passed to the \NAME program via the variable W0\_PH to be read from the input file. The displacement field acting on the phonon is passed thorugh the variable A\_PH from the input file. The electron-phonon coupling matrix $g_{\alpha \beta}$ instead can be passed in two different ways. If the matrix is diagonal ($g_{\alpha \beta} = g_\alpha \delta_{\alpha \beta}$), from the input file one can simply write in order the values $g_\alpha$ in the variable G\_PH. For any general shape one can pass to the input file a GPHfile variable containing the name of the file storing the $g_{\alpha \beta}$ array in Fortran style.
If GPHfile=NONE the matrix $g_{\alpha \beta}$ is read from the variable G\_PH, else it is read from the GPHfile.

In the following we show for a fixed value of $\lambda=2$ the adiabatic to anti-adiabatic crossover for the uniform solution of the Holstein problem.

PLOT docc vs W0, superc and normal
PLT delta vs w0, superc

>>>>

In the following we show for a fixed value of $\lambda/D=1.0$ the adiabatic to anti-adiabatic crossover for the uniform solution of the Holstein problem on the Bethe lattice.

The Holstein interaction takes a simpler form for $\omega_0 \rightarrow \infty$ with $g^2/\omega_0=\lambda/2$ constant, also known as anti-adiabatic limit, for which it becomes a Hubbard attraction:
\begin{equation}
    H_{int} \overset{ \omega_0 \rightarrow \infty}{ \longrightarrow } -\frac{\lambda}{2} \sum_i \Big[\sum_\sigma(c^\dagger_{i\sigma}c_{i\sigma} -\frac{1}{2}) \Big]^2
\end{equation}

Meanwhile, in the opposite limit, $\omega_0 \rightarrow 0$ with $g^2/\omega_0=\lambda/2$ constant, also known as adiabatic limit, the system is in a phase of Bipolaronic insulator~\citep{Capone2006PRB} for $\lambda=1$.

\begin{figure}[ht!]
    \includegraphics[width=\linewidth]{edipack2_examples/Bethe_Holstein/figBethe_Holstein.pdf}
    \caption{\label{figEx5}Plots of the double occupancy (A) and the superconductive order parameter (B) as a function of $\omega_0/D$ for $\lambda/D=1$ in the normal (orange) and superconductive (red) phase. The horizontal broken line are the values for the corresponding Hubbard model in the anti-adiabatic limit.}
\end{figure}
